---
title: "Class 9"
author: "Ralph Goguanco"
output: pdf_document
---

---
title: "Class09"
output: class09_document
---

```{r}
DataExp.df <- read.csv("E:\\Downloads\\DataExportSummary.csv")
DataExp.df

```

> Q1: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

```{r}
(163351+10139)/187423
```
  

> Q2: What proportion of structures in the PDB are protein?

```{r}
163543/187423
```


> Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There are 4496 total structures in the current PDB

> Q4: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?

It would be too crowded if we added three atoms per water molecule since there's a lot of them.

>Q5: There is a conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have (see note below)?

The residue number of the atom associated with the water molecule is ALA28:CA



##3. Introduction to BIO3D in R

```{r}
library(bio3d)

read.pdb(file = "1hsg")
```

>Q7: How many amino acid residues are there in this pdb object? 

There are 198 amino acid residues in this pdb object.

>Q8: Name one of the two non-protein residues? 

Water and MK1

>Q9: How many protein chains are in this structure? 

There are 2 protein chains in this structure

>Q10. Which of the packages above is found only on BioConductor and not CRAN? 

msa

>Q11. Which of the above packages is not found on BioConductor or CRAN?: 


bio3d-view


>Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?

TRUE



```{r}
library(bio3d)
aa <- get.seq("1ake_A")

aa
```

>Q13. How many amino acids are in this sequence, i.e. how long is this sequence? 

214

```{r}
# Blast or hmmer search 
b <- blast.pdb(aa)
```

```{r}
# Plot a summary of search results
hits <- plot(b)
```

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')

```

```{r}
#Download related PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

```{r}
# Align releated PDBs
#pdbs <- pdbaln(files, fit = TRUE), exefile="msa")
```



```{r}
# Vector containing PDB codes for figure axis
#ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
#(pdbs, labels=ids)
```

> What do you note about this plot? Are the black and colored lines similar or different? Where do you think they differ most and why?


The black and colored lines are different, they differe in fluctuation heights but the patterns are almost similar.












