---
title: "Vaccination Mini Project"
author: "Ralph Goguanco"
date: "3/11/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

The goal of this hands-on mini-project is to examine and compare the Covid-19 vaccination rates around San Diego.


```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

>Q1. What column details the total number of people fully vaccinated? 

persons_fully_vaccinated

>Q2. What column details the Zip code tabulation area? 

zip_code_tabulation_area

>Q3. What is the earliest date in this dataset? 

```{r}
head(vax$as_of_date)
#2021-01-05
```

>Q4. What is the latest date in this dataset?

```{r}
tail(vax$as_of_date)
#2022-03-08
```


As we have done previously, let’s call the skim() function from the skimr package to get a quick overview of this dataset:
```{r}
skimr::skim(vax)
```

>Q5. How many numeric columns are in this dataset?

9
>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
N <- sum(is.na(vax$persons_fully_vaccinated))
total <- sum(vax$percent_of_population_fully_vaccinated, na.rm = TRUE)
```

>What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
N/total
```


#Working with dates
```{r}
library(lubridate)
today()

#Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

>Q9. How many days have passed since the last update of the dataset?
3 days

```{r}
today() - vax$as_of_date[109358]
```
>Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?
62

```{r}
date_uniq <- unique(vax$as_of_date)
length(date_uniq)
```

#Working with ZIP codes


```{r}
library(zipcodeR)
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

```{r}
# Subset to San Diego county only areas
sd <- vax[ 4 , ]
```


```{r}
library(dplyr)
sd <- filter(vax, county == "San Diego")

nrow(sd)
```

>Q11. How many distinct zip codes are listed for San Diego County?


```{r}
length(unique(sd))

```

>Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
which.max(sd[, 7])
# Zip Code in #103 is 92154
```


```{r}
sd1 <- filter(sd, as_of_date == "2022-02-22")
```

>Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-02-22”?

```{r}
mean(sd1$percent_of_population_fully_vaccinated, na.rm = TRUE)
```
71.46%

>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-02-22”?

```{r}
hist(sd$percent_of_population_fully_vaccinated)
```


Focus on UCSD/La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population

```

>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
library(ggplot2)
ggplot(ucsd) +
  aes(x=ucsd$as_of_date,
      y = ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")

```

Comparing to similar sized areas

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-02-22")

head(vax.36)
```
>Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
ggplot(ucsd) +
aes(x=ucsd$as_of_date,
      y = ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated") +
  geom_hline(yintercept = 0.732736)

```

>Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”?

```{r}
#fivenum(vax.36)
```

>Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36, aes(x=vax.36$percent_of_population_fully_vaccinated)) + 
geom_histogram()
```

>Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-02-22") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

```
Below the average value

>Q20.  Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
#vax.36.all <- filter(vax, age5_plus_population > 36144)


#ggplot(vax.36.all) +
# aes(x = vax.36$as_of_date,
#      percent_of_population_fully_vaccinated, 
#      group=zip_code_tabulation_area) +
#  geom_line(alpha=0.2, color=vax) +
#  ylim(vax.36) +
# labs(x="Date" , y="Percent",
#       title="Vaccination rate accross California",
#       subtitle="Only areas with a population above 36k are shown") +
#  geom_hline(yintercept = 0.7327, linetype=dash)
```

>Q21. How do you feel about traveling for Spring Break and meeting for in-person class afterwards?

I feel better knowing that the percent vaccinated will only go up



